---
title: "PA1_template"
author: "Biscuit7"
date: "`r format(Sys.Date())`"
output:
      html_document:
            keep_md: TRUE
---

```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE)
library("lubridate")
library("lattice")
```

Analyzing Step Data from an Activity Monitoring Device
======================================================

This project is part of the Reproducible Research course in Coursera's Data Science specialisation track. The purpose of the project is to analyse data collected from a FitBit by way of:

* loading and preprocessing data
* drawing time series plots and histograms
* imputing missing values
* interpreting data to answer research questions

## Loading and preprocessing the data
###Read activity data and remove missing values
```{r}
df <- read.csv("activity.csv",header=T,na.strings = "NA")
df1 <- df[!is.na(df$steps),]
```

## What is mean total number of steps taken per day?
### Histogram 
``` {r}
total <- aggregate(steps ~ date, df1, sum)
hist(total$steps, breaks=20, main="Total Steps Each Day", col="red", xlab="Number of Steps")
```

### Mean and median of the total number of steps taken per day
``` {r}
step_mean <- round(mean(total$steps),2)
step_median <- median(total$steps)
```
The `mean` is `r step_mean` and the `median` is `r step_median`.

## What is the average daily activity pattern?
### Time series of the average steps per day
``` {r}
avg_steps <- aggregate(steps ~ interval, df1, mean)
avg_steps$day <- seq.int(nrow(avg_steps))
plot(avg_steps$interval, avg_steps$steps, type="l", main="Average Steps per Day", col="green", 
     xlab="Interval", ylab="Steps")
```

### The 5-minute interval with the maximum number of steps, on average
``` {r}
avg_int <- aggregate(steps ~ interval, df1, mean)
max_int <- avg_int[avg_int$steps==max(avg_int$steps),"interval"]
```
The `interval` with the maximum number of steps, on average, is `r max_int`.

## Imputing missing values
### Number of records with missing values
``` {r}
sum(is.na(df))
```
There are `r sum(is.na(df))` `missing values`.

### Imputing
Imputation will involve filling in the missing values with the mean number of steps per interval, which is assigned to `avg_int`.

``` {r}
df_new <- df
avg  <- tapply(df_new$steps, df_new$interval, mean, na.rm=T, simplify=T)
df_new$steps[is.na(df_new$steps)] <- avg[as.character(df_new$interval[is.na(df_new$steps)])]
```

### Histogram with imputed values
``` {r}
total_new <- aggregate(steps ~ date, df_new, sum)
hist(total_new$steps, breaks=20, main="Total Steps Each Day", col="blue", xlab="Number of Steps")
```

### Mean and median of the total number of steps taken per day, with imputed values
``` {r}
new_mean <- round(mean(total_new$steps),2)
new_median <- median(total_new$steps)
```
The `new mean` is `r new_mean` and the `new median` is `r new_median`.

## Are there differences in activity patterns between weekdays and weekends?
### Panel plot comparing weekdays and weekends
``` {r}
weekend <- c("Saturday", "Sunday")
df_new$day = as.factor(ifelse(is.element(weekdays(as.Date(df_new$date)),weekend), "Weekend", "Weekdays"))
total_day <- aggregate(steps ~ interval + day, df_new, mean)
xyplot(total_day$steps ~ total_day$interval|total_day$day, col="green", main="Average Steps per Day by Interval", xlab="Interval", ylab="Steps",layout=c(2,1), type="l")
```